\LARGE \bf Geometric Adaptive Control of Attitude Dynamics on $\SO$\\ with State Inequality Constraints

### 2016-02-12 ACC16.2.tex
Add comments to long version from Dr. Lee

The two-sphere is the manifold of unit-vectors in \( \R^3 \) such that \( \Sph^2 = \{ q \in \R^3 \,  \vert \, \norm{q} = 1 \}\).
We define \( r \in \Sph^2 \) to be a unit vector from the mass center of the rigid body along a certain direction and it is represented with respect to the body-fixed frame.
For example, \( r \) may represent the pointing direction of an on-board optical sensor.
We define \( v \in \Sph^2 \) to be a unit vector from the mass center of the rigid body towards an undesired pointing direction and represented in the inertial reference frame.
For example, \( v \) may represent the inertial direction of a bright celestial object or the incoming direction of particles or other debris.
It is further assumed that optical sensor has a strict non-exposure constraint with respect to the celestial object.
We formulate this hard constraint as
\begin{align}
	r^T R^T v \leq \cos \theta , \label{eqn:constraint}
\end{align}
where we assume \( \ang{0} \leq \theta \leq \ang{90}  \) is the required minimum angular separation between \( r \) and \( R^T v \).

@@

To incorporate the state inequality constraints we apply a logarithmic barrier term.
Barrier functions are typically used in optimal control and motion planning applications.
A visualization of the configuration error function is presented in~\cref{fig:avoid_error} which shows that as the boundary of the constraint is neared, or \( r^T R^T v \to \cos \theta \), the barrier term increases, \( B \to \infty\).
We use the scale factor~\(\frac{1}{1+\cos \theta} \) to ensure that \( \Psi \) remains positive definite.
The logarithmic function is popular as it quickly decays away from the constraint boundary.
The positive constant \( \alpha \) serves to shape the barrier function.
As \( \alpha \) is increased the impact of \( B(R) \) is reduced away from the constraint boundary. 
The super position of the attractive and repulsive functions is shown in~\cref{fig:combined_error}.
The control system is defined such that the attitude trajectory follows the negative gradient of \( \Psi \) towards the minimum at \( R = R_d \), while avoiding the constrained region.
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.45\columnwidth} 
		\includegraphics[width=\columnwidth]{attract_error} 
		\caption{Attractive \( A(R) \) } \label{fig:attract_error} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.45\columnwidth} 
		\includegraphics[width=\columnwidth]{avoid_error} 
		\caption{Repulsive \( B(R) \)} \label{fig:avoid_error} 
	\end{subfigure}
	\centering
	\begin{subfigure}[htbp]{0.45\columnwidth} 
		\includegraphics[width=\columnwidth]{combined_error} 
		\caption{Configuration \( \Psi \)} \label{fig:combined_error} 
	\end{subfigure}
	\caption{Configuration error function visualization}
	\label{fig:config_error} 
\end{figure}

@@

This proposition only guarantees that the attitude error vector \( e_R \) asymptotically converges to zero.
However, this does not necessarily imply that \( R \to R_d \) as \( t \to \infty \), since there are three additional critical points of \( \Psi \) where \( e_R = 0 \).
At an undesired equilibrium \( R = \exp{ \parenth{\pi \hat{e}_i}} R_d \) and \( e_\Omega =0 \).
However, we can show that these undesired equilibrium points are unstable in the sense of Lyapunov~\cite{LeeITCST13}.
As a result, we can claim that the desired equilibrium \( R = R_d \text{ and } e_\Omega = 0 \) is almost globally asymptotically stable, which means that the set of initial conditions that do not converge to the desired attitude has zero Lebesgue measure.

@@

The initial state is defined as \(R_0 =  \exp(\ang{225} \times \frac{\pi}{180} \hat{e}_3), \Omega_0 = 0\). 
The desired state is \( R_d = I,\Omega_d = 0\).
\begin{table}%[htbp]
\caption{Constraint Parameters~\label{tab:constraints}}
\begin{center}\begin{tabular}{lc}
Constraint Vector (\( v \)) & Angle (\( \theta \)) \\ \hline \hline 
\([0.174,\,-0.934,\, -0.034]^T\) & \ang{40} \\ \hline 
\([0 ,\, 0.7071 ,\, 0.7071]^T\) & \ang{40} \\ \hline 
\([-0.853 ,\, 0.436 ,\, -0.286]^T\) & \ang{40} \\ \hline 
\([-0.122 ,\,-0.140,\, -0.983]^T\) & \ang{20}\end{tabular} 
\end{center}
\end{table}
We show simulation results for the system stabilizing about the desired attitude with and without the adaptive update law from~\Cref{prop:adaptive_control}.
We assume a fixed disturbance of \(\Delta = \begin{bmatrix} 0.2 & 0.2 & 0.2 \end{bmatrix}^T \), with the function \( W(R,\Omega) = I \).
This form is equivalent to an integral control term which penalizes deviations from the desired configuration.
The first term of~\cref{eqn:delta_dot} has the effect of increasing the proportional gain of the control system, since the time derivative of the attitude error vector, \( \dot{e}_{R} \), is linear with respect to the angular velocity error vector \( e_\Omega\).
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{eR_con} 
		\caption{Attitude error vector \(e_R\) } \label{fig:eR_con} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{Psi_con} 
		\caption{Configuration error \( \Psi \)} \label{fig:Psi_con} 
	\end{subfigure}
	\caption{Attitude stabilization without adaptive update law}
	\label{fig:con} 
\end{figure}
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{eR_adapt} 
		\caption{Attitude error vector \(e_R\) } \label{fig:eR_adapt} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{Psi_adapt} 
		\caption{Configuration error \( \Psi \)} \label{fig:Psi_adapt} 
	\end{subfigure}

@@

Simulation results without the adaptive update law are shown in~\cref{fig:con}.
Without the update law, the system does not achieve zero steady state error. 
\Cref{fig:Psi_con} shows that the configuration error function does not converge to zero and there exist steady state errors.
\Cref{fig:adapt} shows the results with the addition of the adaptive update law.
The addition of the adaptive update law allows the system to converge to the desired attitude in the presence of constraints.
The path of the body fixed sensor in the inertial frame, namely \( R r \), is illustrated in~\cref{fig:cad_adapt}.
The initial attitude is represented with the green circle while the final attitude is marked with a green \(\times\).
The inequality constraints from~\Cref{tab:constraints} are depicted as red cones, where the cone half angle is \( \theta \).
The control system is able to asymptotically converge to zero attitude error.
In addition, the estimate of the disturbance converges to the the true value.

>>>>

The two-sphere is the manifold of unit-vectors in \( \R^3 \) such that \( \Sph^2 = \{ q \in \R^3 \,  \vert \, \norm{q} = 1 \}\).
We define \( r \in \Sph^2 \) to be a unit vector from the mass center of the rigid body along a certain direction and it is represented with respect to the body-fixed frame.
For example, \( r \) may represent the pointing direction of an on-board optical sensor.
We define \( v \in \Sph^2 \) to be a unit vector from the mass center of the rigid body toward an undesired pointing direction and represented in the inertial reference frame.
For example, \( v \) may represent the inertial direction of a bright celestial object or the incoming direction of particles or other debris.
It is further assumed that optical sensor has a strict non-exposure constraint with respect to the celestial object.
We formulate this hard constraint as
\begin{align}
	r^T R^T v \leq \cos \theta , \label{eqn:constraint}
\end{align}
where we assume \( \ang{0} \leq \theta \leq \ang{90}  \) is the required minimum angular separation between \( r \) and \( R^T v \).

@@

To incorporate the state inequality constraints we apply a logarithmic barrier term.
Barrier functions are typically used in optimal control and motion planning applications.
A visualization of the configuration error function is presented in~\cref{fig:avoid_error} which shows that as the boundary of the constraint is neared, or \( r^T R^T v \to \cos \theta \), the barrier term increases, \( B \to \infty\).
We use the scale factor~\(\frac{1}{1+\cos \theta} \) to ensure that \( \Psi \) remains positive definite.
The logarithmic function is popular as it quickly decays away from the constraint boundary.
The positive constant \( \alpha \) serves to shape the barrier function.
As \( \alpha \) is increased the impact of \( B(R) \) is reduced away from the constraint boundary. 
The super position of the attractive and repulsive functions is shown in~\cref{fig:combined_error}.
The control system is defined such that the attitude trajectory follows the negative gradient of \( \Psi \) toward the minimum at \( R = R_d \), while avoiding the constrained region.
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.45\columnwidth} 
		\includegraphics[width=\columnwidth]{attract_error} 
		\caption{Attractive \( A(R) \) } \label{fig:attract_error} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.45\columnwidth} 
		\includegraphics[width=\columnwidth]{avoid_error} 
		\caption{Repulsive \( B(R) \)} \label{fig:avoid_error} 
	\end{subfigure}
	\centering
	\begin{subfigure}[htbp]{0.45\columnwidth} 
		\includegraphics[width=\columnwidth]{combined_error} 
		\caption{Configuration \( \Psi \)} \label{fig:combined_error} 
	\end{subfigure}
	\caption{Configuration error function visualization}
	\label{fig:config_error} 
\end{figure}

@@

This proposition only guarantees that the attitude error vector \( e_R \) asymptotically converges to zero.
However, this does not necessarily imply that \( R \to R_d \) as \( t \to \infty \), since there are at most three additional critical points of \( \Psi \) where \( e_R = 0 \) and \( R^T v = \pm r\).
At an undesired equilibrium \( R = \exp{ \parenth{\pi \hat{e}_i}} R_d \) and \( e_\Omega =0 \).
However, we can show that these undesired equilibrium points are unstable in the sense of Lyapunov~\cite{LeeITCST13}.
As a result, we can claim that the desired equilibrium \( R = R_d \text{ and } e_\Omega = 0 \) is almost globally asymptotically stable, which means that the set of initial conditions that do not converge to the desired attitude has zero Lebesgue measure.

@@

The initial state is defined as \(R_0 =  \exp(\ang{225} \times \frac{\pi}{180} \hat{e}_3), \Omega_0 = 0\). 
The desired state is \( R_d = I,\Omega_d = 0\).
\begin{table}%[htbp]
\caption{Constraint Parameters~\label{tab:constraints}}
\begin{center}\begin{tabular}{lc}
Constraint Vector (\( v \)) & Angle (\( \theta \)) \\ \hline \hline 
\([0.174,\,-0.934,\, -0.034]^T\) & \ang{40} \\ \hline 
\([0 ,\, 0.7071 ,\, 0.7071]^T\) & \ang{40} \\ \hline 
\([-0.853 ,\, 0.436 ,\, -0.286]^T\) & \ang{40} \\ \hline 
\([-0.122 ,\,-0.140,\, -0.983]^T\) & \ang{20}\end{tabular} 
\end{center}
\end{table}
We show simulation results for the system stabilizing about the desired attitude with and without the adaptive update law from~\Cref{prop:adaptive_control}.
We assume a fixed disturbance of \(\Delta = \begin{bmatrix} 0.2 & 0.2 & 0.2 \end{bmatrix}^T \), with the function \( W(R,\Omega) = I \).
This form is equivalent to an integral control term which penalizes deviations from the desired configuration.
The first term of~\cref{eqn:delta_dot} has the effect of increasing the proportional gain of the control system, since the time derivative of the attitude error vector, \( \dot{e}_{R} \), is linear with respect to the angular velocity error vector \( e_\Omega\).
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{eR_con} 
		\caption{Attitude error vector \(e_R\) } \label{fig:eR_con} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{Psi_con} 
		\caption{Configuration error \( \Psi \)} \label{fig:Psi_con} 
	\end{subfigure}
	\caption{Attitude stabilization without adaptive update law}
	\label{fig:con} 
\end{figure}
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{Psi_adapt} 
		\caption{Configuration error \( \Psi \)} \label{fig:Psi_adapt} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{constraint_angles} 
		\caption{Angle to constraints} \label{fig:con_angles} 
	\end{subfigure}

@@

Simulation results without the adaptive update law are shown in~\cref{fig:con}.
Without the update law, the system does not achieve zero steady state error. 
\Cref{fig:Psi_con} shows that the configuration error function does not converge to zero and there exist steady state errors.
\Cref{fig:adapt} shows the results with the addition of the adaptive update law.
The addition of the adaptive update law allows the system to converge to the desired attitude in the presence of constraints.
The path of the body fixed sensor in the inertial frame, namely \( R r \), is illustrated in~\cref{fig:cad_adapt}.
The initial attitude is represented with the green circle while the final attitude is marked with a green \(\times\).
The inequality constraints from~\Cref{tab:constraints} are depicted as red cones, where the cone half angle is \( \theta \).
The control system is able to asymptotically converge to zero attitude error.
\Cref{fig:con_angles} shows that the angle \( \arccos(r^T R^T v_i) \) to each constraint is satisfied for the entire maneuver.
In addition, the estimate of the disturbance converges to the the true value.
